\bf An Examination of Sport Climbing Competition Format and Scoring System

### 2021-08-04 manuscript/Rmd/draft.Rmd
Merge and commit
Table captions

<!-- # https://www.climbing.com/competition/olympics/a-guide-to-the-olympic-climbing-format/ -->

@@

<!-- 
here are other individual competitions consisted of several events combined that do base their scoring on ranks, such as crossfit competitions.  In each event, points are earned based on the competitor's rank in the event based on a scoring table. (https://games.crossfit.com/announcement/12940), and a competitors final score is based on the sum of their scores across all the events.

@@

In this section, we discuss the results of our simulations described in Section \ref{data-and-methods}. We first look at the chances of advancing to the final round for a qualifier, given that 1) they finish first in any of the three disciplines, and 2) they win the first event (which happens to be speed climbing in reality). Our simulation study shows that a climber is almost guaranteed to make the final round if they win at least one of the three climbing concentrations (99.51%, see Table) or if they win the opening event (99.48%, see Table). After winning the first event, a climber is also more likely to finish first overall (30%, see Table) than any other positions in the overall qualification rankings. Therefore, this product of ranks scoring format highly rewards climbers who can get off to a strong start, which is a major strategy climbing athletes should take note of.

@@

cbind(win_any_qual, win_first_qual) %>% 
  kable(col.names = rep(c("Rank", "Count", "Probability", "Cumulative"), 2),
        caption = "title", 
        booktabs = TRUE, 
        position = "H",
        digits = 4) %>% 
  add_header_above(c("Win Any" = 4, "Win First" = 4)) %>%
  kable_styling(latex_options = c("repeat_header", "scale_down"))
```

@@

cbind(win_any_final, win_first_final) %>% 
  kable(col.names = rep(c("Rank", "Count", "Probability", "Cumulative"), 2),
        caption = "title", 
        booktabs = TRUE, 
        position = "H",
        digits = 4) %>% 
  add_header_above(c("Win Any" = 4, "Win First" = 4)) %>%
  kable_styling(latex_options = c("repeat_header", "scale_down"))
```

@@

### Probability of Reordering in Higher Ranks
We first look at the rate at which the ranks above that of which is dropped are reordered.
```{r, echo = FALSE}
load("draft_files/reorder_all_above_probs.rdata")

>>>>

<!-- https://www.climbing.com/competition/olympics/a-guide-to-the-olympic-climbing-format/ -->

@@

<!-- 
There are other individual competitions consisted of several events combined that do base their scoring on ranks, such as crossfit competitions.  In each event, points are earned based on the competitor's rank in the event based on a scoring table. (https://games.crossfit.com/announcement/12940), and a competitors final score is based on the sum of their scores across all the events.

@@

In this section, we discuss the results of our simulations described in Section \ref{data-and-methods}. We first look at the chances of advancing to the final round for a qualifier, given that 1) they finish first in any of the three disciplines, and 2) they win the first event (which happens to be speed climbing in reality). Our simulation study shows that a climber is almost guaranteed to make the final round if they win at least one of the three climbing concentrations or if they win the opening event, as the probabilities of advancing for these two situations are 0.9948 and 0.9951, respectively (see Table 2). Moreover, after winning the first event, a climber is also more likely to finish first overall (see Table 2) than any other positions in the overall qualification rankings. Therefore, this product of ranks scoring format highly rewards climbers who can get off to a strong start, which is a major strategy climbing athletes should take note of.

@@

cbind(win_any_qual, win_first_qual) %>% 
  kable(col.names = rep(c("Rank", "Count", "Probability", "Cumulative"), 2),
        caption = "This table shows every possible qualification rank obtained from simulations for a climber given that a they win at least one discipline or finish first in the first discipline. For each rank, information on the number of times that climbers finish at the given rank, the probability of finishing at exactly the given rank, and the probability of finishing at or below the given rank are included.", 
        booktabs = TRUE, 
        position = "H",
        digits = 4) %>% 
  add_header_above(c("Win Any" = 4, "Win First" = 4)) %>%
  kable_styling(latex_options = c("repeat_header", "scale_down"))
```

@@

cbind(win_any_final, win_first_final) %>% 
  kable(col.names = rep(c("Rank", "Count", "Probability", "Cumulative"), 2),
        caption = "This table shows every possible final rank obtained from simulations for a climber given that a they win at least one discipline or finish first in the first discipline. For each rank, information on the number of times that climbers finish at the given rank, the probability of finishing at exactly the given rank, and the probability of finishing at or below the given rank are included.", 
        booktabs = TRUE, 
        position = "H",
        digits = 4) %>% 
  add_header_above(c("Win Any" = 4, "Win First" = 4)) %>%
  kable_styling(latex_options = c("repeat_header", "scale_down"))
```

@@

### Probability of Reordering in Higher Ranks

We first look at the rate at which the ranks above that of which is dropped are reordered.

```{r, echo = FALSE}
load("draft_files/reorder_all_above_probs.rdata")

### 2021-08-04 manuscript/Rmd/draft.tex
Table captions

In this section, we discuss the results of our simulations described in
Section \ref{data-and-methods}. We first look at the chances of
advancing to the final round for a qualifier, given that 1) they finish
first in any of the three disciplines, and 2) they win the first event
(which happens to be speed climbing in reality). Our simulation study
shows that a climber is almost guaranteed to make the final round if
they win at least one of the three climbing concentrations (99.51\%, see
Table) or if they win the opening event (99.48\%, see Table). After
winning the first event, a climber is also more likely to finish first
overall (30\%, see Table) than any other positions in the overall
qualification rankings. Therefore, this product of ranks scoring format
highly rewards climbers who can get off to a strong start, which is a
major strategy climbing athletes should take note of.

@@

\caption{\label{tab:unnamed-chunk-6}title}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{rrrrrrrr}
\toprule
\multicolumn{4}{c}{Win Any} & \multicolumn{4}{c}{Win First} \\
\cmidrule(l{3pt}r{3pt}){1-4} \cmidrule(l{3pt}r{3pt}){5-8}
Rank & Count & Probability & Cumulative & Rank & Count & Probability & Cumulative\\
\midrule
1 & 7890 & 0.2762 & 0.2762 & 1 & 2999 & 0.2999 & 0.2999\\
2 & 5943 & 0.2080 & 0.4842 & 2 & 1963 & 0.1963 & 0.4962\\
3 & 4644 & 0.1626 & 0.6467 & 3 & 1559 & 0.1559 & 0.6521\\
4 & 3651 & 0.1278 & 0.7745 & 4 & 1269 & 0.1269 & 0.7790\\
5 & 2766 & 0.0968 & 0.8714 & 5 & 957 & 0.0957 & 0.8747\\
\addlinespace
6 & 1960 & 0.0686 & 0.9400 & 6 & 686 & 0.0686 & 0.9433\\
7 & 1118 & 0.0391 & 0.9791 & 7 & 363 & 0.0363 & 0.9796\\
8 & 448 & 0.0157 & 0.9948 & 8 & 155 & 0.0155 & 0.9951\\
9 & 127 & 0.0044 & 0.9992 & 9 & 42 & 0.0042 & 0.9993\\
10 & 19 & 0.0007 & 0.9999 & 10 & 6 & 0.0006 & 0.9999\\
\addlinespace
11 & 3 & 0.0001 & 1.0000 & 11 & 1 & 0.0001 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-9}title}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{rrrrrrrr}
\toprule
\multicolumn{4}{c}{Win Any} & \multicolumn{4}{c}{Win First} \\
\cmidrule(l{3pt}r{3pt}){1-4} \cmidrule(l{3pt}r{3pt}){5-8}
Rank & Count & Probability & Cumulative & Rank & Count & Probability & Cumulative\\
\midrule
1 & 9277 & 0.3512 & 0.3512 & 1 & 4295 & 0.4295 & 0.4295\\
2 & 7409 & 0.2805 & 0.6317 & 2 & 2505 & 0.2505 & 0.6800\\
3 & 5257 & 0.1990 & 0.8308 & 3 & 1702 & 0.1702 & 0.8502\\
4 & 3078 & 0.1165 & 0.9473 & 4 & 1020 & 0.1020 & 0.9522\\
5 & 1187 & 0.0449 & 0.9922 & 5 & 393 & 0.0393 & 0.9915\\
\addlinespace
6 & 198 & 0.0075 & 0.9997 & 6 & 83 & 0.0083 & 0.9998\\
7 & 7 & 0.0003 & 1.0000 & 7 & 2 & 0.0002 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

>>>>

In this section, we discuss the results of our simulations described in
Section \ref{data-and-methods}. We first look at the chances of
advancing to the final round for a qualifier, given that 1) they finish
first in any of the three disciplines, and 2) they win the first event
(which happens to be speed climbing in reality). Our simulation study
shows that a climber is almost guaranteed to make the final round if
they win at least one of the three climbing concentrations or if they
win the opening event, as the probabilities of advancing for these two
situations are 0.9948 and 0.9951, respectively (see Table 2). Moreover,
after winning the first event, a climber is also more likely to finish
first overall (see Table 2) than any other positions in the overall
qualification rankings. Therefore, this product of ranks scoring format
highly rewards climbers who can get off to a strong start, which is a
major strategy climbing athletes should take note of.

@@

\caption{\label{tab:unnamed-chunk-6}This table shows every possible qualification rank obtained from simulations for a climber given that a they win at least one discipline or finish first in the first discipline. For each rank, information on the number of times that climbers finish at the given rank, the probability of finishing at exactly the given rank, and the probability of finishing at or below the given rank are included.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{rrrrrrrr}
\toprule
\multicolumn{4}{c}{Win Any} & \multicolumn{4}{c}{Win First} \\
\cmidrule(l{3pt}r{3pt}){1-4} \cmidrule(l{3pt}r{3pt}){5-8}
Rank & Count & Probability & Cumulative & Rank & Count & Probability & Cumulative\\
\midrule
1 & 7890 & 0.2762 & 0.2762 & 1 & 2999 & 0.2999 & 0.2999\\
2 & 5943 & 0.2080 & 0.4842 & 2 & 1963 & 0.1963 & 0.4962\\
3 & 4644 & 0.1626 & 0.6467 & 3 & 1559 & 0.1559 & 0.6521\\
4 & 3651 & 0.1278 & 0.7745 & 4 & 1269 & 0.1269 & 0.7790\\
5 & 2766 & 0.0968 & 0.8714 & 5 & 957 & 0.0957 & 0.8747\\
\addlinespace
6 & 1960 & 0.0686 & 0.9400 & 6 & 686 & 0.0686 & 0.9433\\
7 & 1118 & 0.0391 & 0.9791 & 7 & 363 & 0.0363 & 0.9796\\
8 & 448 & 0.0157 & 0.9948 & 8 & 155 & 0.0155 & 0.9951\\
9 & 127 & 0.0044 & 0.9992 & 9 & 42 & 0.0042 & 0.9993\\
10 & 19 & 0.0007 & 0.9999 & 10 & 6 & 0.0006 & 0.9999\\
\addlinespace
11 & 3 & 0.0001 & 1.0000 & 11 & 1 & 0.0001 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-9}This table shows every possible final rank obtained from simulations for a climber given that a they win at least one discipline or finish first in the first discipline. For each rank, information on the number of times that climbers finish at the given rank, the probability of finishing at exactly the given rank, and the probability of finishing at or below the given rank are included.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{rrrrrrrr}
\toprule
\multicolumn{4}{c}{Win Any} & \multicolumn{4}{c}{Win First} \\
\cmidrule(l{3pt}r{3pt}){1-4} \cmidrule(l{3pt}r{3pt}){5-8}
Rank & Count & Probability & Cumulative & Rank & Count & Probability & Cumulative\\
\midrule
1 & 9277 & 0.3512 & 0.3512 & 1 & 4295 & 0.4295 & 0.4295\\
2 & 7409 & 0.2805 & 0.6317 & 2 & 2505 & 0.2505 & 0.6800\\
3 & 5257 & 0.1990 & 0.8308 & 3 & 1702 & 0.1702 & 0.8502\\
4 & 3078 & 0.1165 & 0.9473 & 4 & 1020 & 0.1020 & 0.9522\\
5 & 1187 & 0.0449 & 0.9922 & 5 & 393 & 0.0393 & 0.9915\\
\addlinespace
6 & 198 & 0.0075 & 0.9997 & 6 & 83 & 0.0083 & 0.9998\\
7 & 7 & 0.0003 & 1.0000 & 7 & 2 & 0.0002 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}
